---
title: "with()"
description: |
  A brief introduction to the with() function
author:
  - name: Kaija Gahm
    url: {}
date: 07-20-2018
output:
  distill::distill_article:
    self_contained: false
    highlight: kate
preview: noPreview.png
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem

Making graphics with base R is annoying for many reasons, but a big one is having to type the name of the data frame over and over again to reference different columns.

### Context

Back to our Mississippi River fish data. I've aggregated my sampling points into polygons, and now I want to explore some of their characteristics. To do that, I'd like to make some tables and plots, and because these are just quick, exploratory plots, I don't feel like dealing with `ggplot`.

Load in the data (accessible on GitHub).

```{r, preview = F}
# Load data from GitHub
polygon <- read.csv("https://tinyurl.com/rta6hkbo") 

# Look at what we're dealing with
dim(polygon) # How big is the data set? 
head(polygon, 3) # Look at the first few rows
```

First, I'd like to see how total volume `tot_vol` of the aquatic area scales with its `Area`.

In base R:

```{r eval = F, preview = F}
# Formula notation
plot(polygon$tot_vol ~ polygon$Area)
# OR: 

# Comma notation
plot(polygon$Area, polygon$tot_vol)
```

Either way, we get this:

```{r echo = F, preview = F}
plot(polygon$Area, polygon$tot_vol)
```

Or a more informative plot, with both variables on a log scale:

```{r, preview = F}
plot(log(polygon$tot_vol) ~ log(polygon$Area))
```

This isn't too too clunky, but if the data frame name or column names are long, it can get a little annoying.

### Solution

The `with()` function allows you to specify the data frame your variables are coming from and then reference the variables with respect to the data frame, similar to the `ggplot` argument `data =`. Handy.

```{r, preview = F}
# Plot using the with() function 
with(polygon, plot(tot_vol ~ Area))
```

You can add any other arguments inside of the function, as normal, it's just now wrapped in `with()`.

```{r, preview = F}
# Log-transform variables and make the points blue dots, because why not?
with(polygon, plot(log(tot_vol) ~ log(Area), # log-transform 
                   pch = 20, # dots instead of circles 
                   col = "blue", # make the dots blue 
                   main = "Polygon volume by area, log-transformed") # title 
     )
```

It's worth noting that this works for other functions besides `plot()`, too. Here's an example with `table()`: let's look at how many sampling polygons include revetment, broken down by navigation pool (area of the river). The data set contains three navigation pools: 4, 8, and 13.

```{r, preview = F}
# Two-way table by pool and revetment
with(polygon, table(revetment, pool))
```

### Outcome

Quick plots and data manipulation made even quicker!

### Resources

Discussion of when to use `with()`:
https://stackoverflow.com/questions/42283479/when-to-use-with-function-and-why-is-it-good

